Aspects of Dealing with Imperfect Data in Temporal Databases

### 2013-02-15 trunk/book/sources/fuzzyTimeDomain.tex
update!

\begin{itemize}
	\item \emph{Uncertainty}: Temporal information or data may contain uncertainty. This means that the exact temporal value is (partially) unknown, however, generally some knowledge is present anyhow, possibly describing the value~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},~\cite{nagypal2003},~\cite{Dubois89},~\cite{343607}. E.g., the temporal notion described in a sentence like `The Belfry of Bruges was finished on a day somewhere between 01/01/1201 A.D. and 31/12/1300 A.D.' contains uncertainty: it is known that the belfry of Bruges was finished on a single day and that this day lays somewhere between 01/01/1201 A.D. and 31/12/1300 A.D., but it is not known exactly which day it is.
	\item \emph{Vagueness}: Temporal information or data might contain inherent vagueness, as a precise instant or time interval may be intended, but the description of it is certainly vague~\cite{schockaert08},~\cite{nagypal2003},~\cite{Dev98}. E.g., the temporal notion described in a sentence like `It happened during summer.' contains vagueness, as even the boundaries of the mentioned temporal notion are not clearly expressed.
	\item \emph{Subjectivity or ambiguity}: Temporal notions might be subject to subjectivity or ambiguity. In certain cases, the temporal notion concerns a historical period like `late romanticism' or `the early middle ages' and thus contains subjectivity~\cite{nagypal2003}. In other cases, the interpretation of the temporal notion depends on extra factors. E.g., consider a person saying to another person `Let's meet each other at six.' The person hearing these words doesn't now if 6 a.m. or 6 p.m. is intended, though the person saying the words does.
\end{itemize}

@@

\item \emph{Indefinite temporal information}~\cite{Dey1996}:
Indefinite temporal information contains information describing a situation in which the time indication associated with some fact has not been fully defined. E.g., consider an event that in fact occured but it is not known exactly when.

\item \emph{Infinite temporal uncertain information}~\cite{Kabanza1990}:
Infinite temporal uncertain information contains information describing a situation in which an infinite number of time indications are associated with some fact. This is usually found in recurrent events like meetings. E.g., consider meetings that take place every Wednesday at noon. Some systems (usually with different information providers) may dispute the occurence and/or the duration of a fact.
\end{itemize}

@@

\begin{example}
Consider a speaker and a hearer. The speaker wants to make an appointment with the hearer. Now, consider the speaker saysing: \\
\begin{center}
\emph{`We will meet each other tomorrow around 10'}\\
\end{center}
The hearer will now usually instinctively agree that the appointment will be in e.g. the time interval between 9.55h and 10.05h.
\end{example}

@@

\begin{itemize}
\item \emph{Probability theory}:
Probability theory~\cite{Dey1996},~\cite{Lakshmanan1997},~\cite{Haddawy1996} is usually employed when uncertainty concerning a time interval allows a probability to be associated to the time interval. The use of probability theory is very usual in logistics information systems. E.g., \emph{`The package will arrive at its destination on monday morning with a probability of 0.8'}.

\item \emph{Possibility theory}:
Using possibility theory~\cite{DidierDubois1988a}, a possibility degree is associated to the temporal fact or event. Possibility theory is widely used to model uncertainty and vaguenes in time~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},\cite{Dubois89},\cite{devos94},\cite{nagypal2003}. Several works~\cite{schockaert08},~\cite{ohlbach2004} present fuzzy versions of the temporal relations proposed by Allen~\cite{Allen83}. The aim of these works is generally to obtain a flexible way to compare uncertain, ill-known temporal intervals by means of temporal relationships. The study of imperfect temporal metadata is done in \cite{Bordogna2009}, \cite{Bordogna2011}. In ~\cite{Soysangwarn2009} a proposal to use in fuzzy databases temporal fuzzy linguistic terms is studied. Burney ~\cite{burney2010a}, \cite{burney2010b} has studied recently the combination of fuzzy databases with temporal data.

>>>>

\begin{itemize}
	\item \emph{Uncertainty}: Temporal information or data may contain uncertainty. This means that the exact temporal value is (partially) unknown, however, generally some knowledge is present anyhow, possibly describing the value~\cite{Dubois2003},~\cite{nagypal2003},~\cite{Dubois1989},~\cite{343607}. E.g., the temporal notion described in a sentence like `The Belfry of Bruges was finished on a day somewhere between 01/01/1201 A.D. and 31/12/1300 A.D.' contains uncertainty: it is known that the belfry of Bruges was finished on a single day and that this day lays somewhere between 01/01/1201 A.D. and 31/12/1300 A.D., but it is not known exactly which day it is.
	\item \emph{Vagueness}: Temporal information or data might contain inherent vagueness, as a precise instant or time interval may be intended, but the description of it is certainly vague~\cite{schockaert08},~\cite{nagypal2003},~\cite{Dev98}. E.g., the temporal notion described in a sentence like `It happened during summer.' contains vagueness, as even the boundaries of the mentioned temporal notion are not clearly expressed.
	\item \emph{Subjectivity or ambiguity}: Temporal notions might be subject to subjectivity or ambiguity. In certain cases, the temporal notion concerns a historical period like `late romanticism' or `the early middle ages' and thus contains subjectivity~\cite{nagypal2003}. In other cases, the interpretation of the temporal notion depends on extra factors. E.g., consider a person saying to another person `Let's meet each other at six.' The person hearing these words doesn't now if 6 a.m. or 6 p.m. is intended, though the person saying the words does.
\end{itemize}

@@

\item \emph{Indefinite temporal information}~\cite{Dey1996}:
Indefinite temporal information contains information describing a situation in which the time indication associated with some fact has not been fully defined. E.g., consider an event that in fact occurred but it is not known exactly when.

\item \emph{Infinite temporal uncertain information}~\cite{Kabanza1990}:
Infinite temporal uncertain information contains information describing a situation in which an infinite number of time indications are associated with some fact. This is usually found in recurrent events like meetings. E.g., consider meetings that take place every Wednesday at noon. Some systems (usually with different information providers) may dispute the occurrence and/or the duration of a fact.
\end{itemize}

@@

\begin{example}
Consider a speaker and a hearer. The speaker wants to make an appointment with the hearer. Now, consider the speaker saying: \\
\begin{center}
\emph{`We will meet each other tomorrow around 10'}\\
\end{center}
The hearer will now usually instinctively agree that the appointment will be in e.g. the time interval between 9.55h and 10.05h.
\end{example}

@@

\begin{itemize}
\item \emph{Probability theory}:
Probability theory~\cite{Dey1996},~\cite{Lakshmanan1997},~\cite{Haddawy1996} is usually employed when uncertainty concerning a time interval allows a probability to be associated to the time interval. The use of probability theory is very usual in logistics information systems. E.g., \emph{`The package will arrive at its destination on Monday morning with a probability of 0.8'}.

\item \emph{Possibility theory}:
Using possibility theory~\cite{DidierDubois1988a}, a possibility degree is associated to the temporal fact or event. Possibility theory is widely used to model uncertainty and vagueness in time~\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in},\cite{Dubois89},\cite{devos94},\cite{nagypal2003}. Several works~\cite{schockaert08},~\cite{ohlbach2004} present fuzzy versions of the temporal relations proposed by Allen~\cite{Allen83}. The aim of these works is generally to obtain a flexible way to compare uncertain, ill-known temporal intervals by means of temporal relationships. The study of imperfect temporal metadata is done in \cite{Bordogna2009}, \cite{Bordogna2011}. In ~\cite{Soysangwarn2009} a proposal to use in fuzzy databases temporal fuzzy linguistic terms is studied. Burney ~\cite{burney2010a}, \cite{burney2010b} has studied recently the combination of fuzzy databases with temporal data.

### 2013-02-15 trunk/book/sources/intro.tex
update!

Humans handle temporal information using certain temporal notions like time intervals or time points~\cite{Dyreson1994}, and they often have to deal with imperfections like imprecisions, vaguenesses, uncertainties or inconsistencies possibly contained in the descriptions of these temporal notions. Among many others, these possible imperfections in descriptions of temporal notions determine an important issue in temporal modelling. e.g., the description of the temporal notion in a sentence like `The Belfry of Bruges was finished on a day somewhere between 01/01/1201 A.D. and 31/12/1300 A.D.' contains imperfection because of the uncertainty in the used time-related expression. It is known that the building was finished on a single day, but it is not known precisely which day this was.

@@

The aim of this chapter is to present and explain some main concepts regarding time in information systems and to present and discuss some issues and techniques concerned with handling data imperfections related to time. The rest of this chapter is structured as follows. Section \ref{sec:time-domain} presents some basic concepts and terminology about temporal modelling and discusses some of its important aspects and issues, while section~\ref{sec:fuzzy-time-mod} presents some important issues concerning the combination of data imperfections and temporal modelling. In section \ref{sec:temporal-databases}, some basic concepts and terminology about temporal databases are presented, followed by an overview of some interesting issues concerning temporal databases and a survey of some commercial temporal database systems. Finally, in section \ref{sec:imprecision-uncertainty}, an approach to querying temporal databases containing imperfect temporal information is presented, followed by some conclusions and some suggestions for future work in section \ref{sec:further-research-work}.

>>>>

Humans handle temporal information using certain temporal notions like time intervals or time points~\cite{Dyreson1994}, and they often have to deal with imperfections like imprecision's, vagueness, uncertainties or inconsistencies possibly contained in the descriptions of these temporal notions. Among many others, these possible imperfections in descriptions of temporal notions determine an important issue in temporal modelling. e.g., the description of the temporal notion in a sentence like `The Belfry of Bruges was finished on a day somewhere between 01/01/1201 A.D. and 31/12/1300 A.D.' contains imperfection because of the uncertainty in the used time-related expression. It is known that the building was finished on a single day, but it is not known precisely which day this was.

@@

The aim of this chapter is to present and explain some main concepts regarding time in information systems and to present and discuss some issues and techniques concerned with handling data imperfections related to time. The rest of this chapter is structured as follows. Section \ref{sec:time-domain} presents some basic concepts and terminology about temporal modelling and discusses some of its important aspects and issues, while section~\ref{sec:fuzzy-time-mod} presents some important issues concerning the combination of data imperfections and temporal modelling. In section~\ref{sec:temporal-databases}, some basic concepts and terminology about temporal databases are presented, followed by an overview of some interesting issues concerning temporal databases and a survey of some commercial temporal database systems. Finally, in section~\ref{sec:imprecision-uncertainty}, an approach to querying temporal databases containing imperfect temporal information is presented, followed by some conclusions and some suggestions for future work in section~\ref{sec:further-research-work}.
